---
title: "Building a cross-device dataset"
author: "Trevor Paulsen"
output: html_notebook
---

Cross device analysis is a really tricky problem! That's because there's a lot of IDs floating around and you need someway to make sense of all of them. Typically you'll have multiple cookie IDs for a single user ID (users with multiple devices), and also multiple users on a single cookie (shared devices).

Do identity map, you'll need either 1) a user/login ID stored in your data like a prop or eVar or 2) an unobfuscated IP address (some localities aren't allowed to store this, so be cautious)

Also - identity mapping can pose a privacy risk depending on your area of business and the countries in which you operate. I'm not a lawyer, so if you're not 100% sure on what's ok and not ok for your business, best to check with your own attorney!

The basic strategy is:
  1) combine all of the cookie/device IDs from the same IP address
  2) split, rejoin, and merge all of those device+IP IDs based on the user IDs
  
More details: http://datafeedtoolbox.com/attribution-theory-cross-device-marketing-attribution-with-apache-spark-and-r/

If you don't want to use the IP address at all, just skip to step 2

```{r}
# Step 1 - Combining Devices Using IP Address
# Step 1a - identify which IPs are good for stitching

ips_to_use = data_feed_ready %>%
  group_by(ip) %>%
  summarise(
    visitors = n_distinct(visitor_id)
  ) %>%
  # Useful IPs happen on more than one cookie, but I'll also
  # throw out IPs that have more than 20 cookies to prevent
  # over-stitching (like a whole office building or something).
  filter(visitors <=20 & visitors > 1) %>%
  collect()

# Plot a histogram of cookies per ip address
library(plotly)
p = plot_ly(x = ips_to_use$visitors, type="histogram") %>%
  layout(
    title="Cookies Per IP Address", 
    xaxis=list(title="Number of IPs"), 
    yaxis=list(title="Number of cookies")
  )
p
```

# Combining cookies based on common IP address

Now we need to identify all of the visitor IDs (cookies) that these IP addresses apply to. Tricky because cookies can have multiple IP addresses (say if you move your device form work to home), but we'll just use the last known IP to simplify things a bit:


```{r}
# Step 1b - map of visitor_ids to its last known IP
visitor_ip_lookup = data_feed_ready %>%
  group_by(visitor_id) %>%
  arrange(hit_time_gmt) %>%
  summarise(
    last_known_ip = last(ip)
    
    # If using Adobe Query Service or Spark SQL, you'd do it like:
    # last_known_ip = last_value(ip)
    
  ) %>%
  ungroup() %>%
  left_join(ips_to_use, by=c("last_known_ip"="ip")) %>%
  mutate(
    
    # this adds a new column for just those IPs that had 2 to 20 visitors
    ip_stitched_user_id = ifelse(is.na(visitors), visitor_id, last_known_ip)
  )
  
# Step 1c - apply the "ip_stitched_user_id" to the data feed
data_feed_ip_blended = data_feed_ready %>%
  left_join(visitor_ip_lookup, by="visitor_id")
  
# Plot a bar chart showing the visitor "compression"
total_visitors_vs_new_ids = data_feed_ip_blended %>%
  summarise(
    total_visitors = n_distinct(visitor_id),
    total_ip_stitched_people = n_distinct(ip_stitched_user_id)
  ) %>% collect()

# Plot the compression
p = plot_ly(
  total_visitors_vs_new_ids, 
  values = c(total_visitors_vs_new_ids$total_visitors, total_visitors_vs_new_ids$total_visitors-total_visitors_vs_new_ids$total_ip_stitched_people),
  type = "pie",
  labels = c("People", "Cookies Compressed")
)
p
```

# Splitting cookies with multiple user IDs

Now that we have stitched our cookies based on IP, we're ready to stitch based on our login ID, or if you skipped that part, you can start here. Stitching using a user ID can be tricky, espeically on a shared device - I don't want to attribute purchases I make to marketing that was exposed to someone else in my family using a shared device. To split a single cookie across multiple users I'm going to use something called "windowing" to create a new column called "id_group" using my customer id "user_id".


```{r}
# Step 2 - stitch the data based on the user ID
# Step 2a: split up cookie IDs with multiple user_ids

data_feed_with_id_group = data_feed_ip_blended %>% # or just use data_feed_ready
  group_by(ip_stitched_user_id) %>% # or just visitor_id
  arrange(hit_time_gmt) %>%
  mutate(
    
    # first, add a 1 everytime I see a user_id
    user_id_present = ifelse(!is.na(user_id), 1, NA),
    
    # then, do a cumsum of that new column to group hits
    group_following_hits = cumsum(ifelse(is.na(user_id_present), 0, user_id_present)),
    
    # finally, bring in hits prior to first user_id
    id_group = ifelse(group_following_hits == 0, 1, group_following_hits)
    
  ) %>%
  ungroup()

# validate with an id that had multiple user_ids 
# ip_address2509 ip_address9963
data_feed_with_id_group %>% filter(ip_stitched_user_id == "ip_address2509") %>%
  select(ip_stitched_user_id, user_id, user_id_present, group_following_hits, id_group)
```

# Combining cookies with common user IDs

Now that we've added a grouping field to split cookies with multiple user_ids, we're ready to create the ID lookup table we'll use for stitching.

```{r}
# Step 2b - create an ID lookup table

id_lookup_table = data_feed_with_id_group %>%
  group_by(ip_stitched_user_id, id_group) %>%
  summarize(
    stitched_id = paste0(user_id[which(!is.na(user_id))],collapse = "")
    
    # for Spark SQL or PostgreSQL:
    # stitched_id = concat_ws("", collect_list(user_id))
    
  ) %>% ungroup() %>%
  group_by(ip_stitched_user_id) %>%
  arrange(id_group) %>%
  mutate(
    lagged_stitched_id = lag(stitched_id),
    stitched_id = ifelse((stitched_id == "") & 
      (id_group == max(id_group)), lagged_stitched_id, stitched_id)
  ) %>%
  mutate(
    stitched_id = ifelse(is.na(stitched_id), ip_stitched_user_id, stitched_id)
  ) %>%
  select(
    ip_stitched_user_id,
    id_group,
    stitched_id
  )


id_lookup_table %>% filter(ip_stitched_user_id == "ip_address2509") 
```

# Doing the final join

Finally, we can join our ID lookup table to the data_feed to give us our new cross-device, stitched user ID for analysis!
```{r}
# Step 2c - joining the ID lookup to the data feed

stitched_data_feed = data_feed_with_id_group %>%
  left_join(id_lookup_table, by=c("ip_stitched_user_id", "id_group"))
  
stitched_data_feed %>% filter(ip_stitched_user_id == "ip_address2509") %>% select(ip_stitched_user_id, ip, user_id, ip_stitched_user_id)
```




#Visualization of the results
Visualizing the compression we just achieved, a 24% overall compression - not too bad!

```{r}
summary_stats = stitched_data_feed %>%
  summarize(
    cookie_count = n_distinct(visitor_id),
    person_count = n_distinct(stitched_id)
  ) %>% collect()

p = plot_ly(
  summary_stats, 
  values = c(summary_stats$person_count, summary_stats$cookie_count-summary_stats$person_count),
  type = "pie",
  labels = c("People", "Cookies Compressed")
)
p
```

