---
title: "Visualizing the Customer Journey"
author: "Trevor Paulsen"
output: html_notebook
---

Next we'll visualize the individual paths users are taking through each of the channels before conversion in two different ways - using a scatter plot as well as a sankey "flow" diagram. For this step we'll use the same "channel stack" concept we used for attribution previously, but this time we'll discard the non-onverting paths and create a new field for conversion rate.

More info: http://datafeedtoolbox.com/visualizing-the-customer-journey-with-r-and-adobe-analytics-data-feeds/

```{r}
# Create a modified channel stacks data frame

channel_stacks = data_feed_sequenced %>%
  group_by(stitched_id, order_seq) %>%
  
  #first remove irrelevant hits:
  filter(!is.na(non_attributed_channel) | orders>0) %>%
  
  #next remove repeated values with a lag function:
  filter((non_attributed_channel != lag(non_attributed_channel, default="1")) | orders>0) %>%
  
  #now concatenate the sequence into a single row:
  summarize(
    path = paste(non_attributed_channel[which(!is.na(non_attributed_channel))], collapse=">"),
    
    # for Spark SQL or PostgreSQL:
    # path = concat_ws(" > ", collect_list(non_attributed_channel))
    
    orders = sum(orders)
  ) %>% ungroup() %>%
  
  #next roll up each unique path by count and conversion:
  group_by(path) %>%
  summarize(
    orders = sum(orders),
    path_count = n()
  ) %>% ungroup() %>%
  
  #last create a conversion rate column and pull it out of Spark:
  mutate(
    conversion_rate = orders/path_count
  ) %>%
  filter(path != "") %>%
  collect()

channel_stacks
```

# Visualizing Paths with a Scatter Plot

Plotly is a great visualization tool - it works great with R, Python, and other languages and the visualizations it produces are highly interactive and sharable. Here, we'll visualize the channel stacks with a cool scatter plot:

```{r message=FALSE}
# Visualizing paths with a scatter plot
p = channel_stacks %>% plot_ly() %>% 
  add_markers(
    y = ~orders, 
    x = ~path_count,
    color = ~conversion_rate, 
    size = ~conversion_rate,
    text = ~path,
    name = "Path"
  ) %>%
  layout(
    xaxis = list(type="log", title="Number of Paths"),
    yaxis = list(type="log", title="Number of Conversions")
  ) %>% 
  colorbar(
    title = "Rate"
  )
p
```

# Visualizing the Path with a Sankey Diagram

The next plot will illustrate how to represent all of these stacked channels into a cool Sankey diagram. The way the Sankey diagram works is to create an array for each "node" denoting how many users went into and out of that specific node. So first step, we'll create a node for each "step" deep we want to go for each channel:


```{r message=FALSE}
# Visualizing customer paths with a Sankey Diagram

# Creating a list of channels for convinience
channel_stacks$path_list = strsplit(x=channel_stacks$path,split=">")

# set the depth of the Sankey Diagram
depth = 4

#Generate node labels and label length vectors
node_labels=rep(list(list()),depth)
label_length = list()
for(i in 1:depth){
  for(j in 1:length(channel_stacks$path)){
    if(!is.na(channel_stacks$path_list[j][[1]][i]))
      node_labels[[i]][j] = channel_stacks$path_list[j][[1]][i]
  }
  node_labels[[i]] = unique(unlist(node_labels[[i]]))
  node_labels[[i]] = node_labels[[i]][order(node_labels[[i]])]
  label_length[[i]] = length(node_labels[[i]])
}
node_labels = unlist(node_labels)
label_length = unlist(label_length)

node_labels
label_length
```

# Building Sankey Nodes

Once I've created those nodes, I'll have to run a bit of code to figure out how many users went to and from each of the nodes. This part is a little messy (I don't typically like to use for loops in my R code), but it works. This gives me a table for each node, the step its in, how many users are going into it, and how many users are going out of it.

```{r message=FALSE}
# Build a data frame to fill out with each path view
combos = NULL
for(i in 1:(depth-1)){
  for(j in (1 + sum(label_length[1:i-1])):(label_length[i] + sum(label_length[1:i-1]))){
    for(k in (1 + label_length[i] + sum(label_length[1:i-1])):(label_length[i+1] + label_length[i] + sum(label_length[1:i-1]))){
      combos = rbind(combos, c(i,j,k,0))
    } 
  }
}
combos = as.data.frame(combos)
names(combos) = c("step","source","target","value")

#Populate the combo table
for(i in 1:(dim(combos)[1])){
  for(j in 1:(dim(channel_stacks)[1])){
    combos$value[i] = sum(combos$value[i], ifelse(
      (node_labels[combos$source[i]] == channel_stacks$path_list[j][[1]][combos$step[i]]) &
      (node_labels[combos$target[i]] == channel_stacks$path_list[j][[1]][combos$step[i]+1]),
      channel_stacks$path_count[j],0), na.rm = TRUE)
  }
}

combos
```

# Adding a Conversion Node

Lastly, I'll create a node that represents a conversion. That'll allow us to visualize how many users actually ended up in a conversion at the end of their journey.

```{r message=FALSE}
#Add a node to populate with conversion values
uniques = unique(c(combos$source,combos$target))
converts = as.data.frame(list("step"=rep(0,length(uniques)), "source"=uniques, "target"=rep(max(uniques)+1,length(uniques)), "value"=rep(0,length(uniques))))
combos = rbind(combos,converts)
for(i in 1:(dim(channel_stacks)[1])){
  stack_depth = min(depth,length(channel_stacks$path_list[i][[1]]))
  index_val = which(combos$step==0 & combos$source==(which(node_labels == channel_stacks$path_list[i][[1]][stack_depth]) + ifelse(stack_depth>1, sum(label_length[1:(stack_depth-1)]),0)))
  combos$value[index_val] = combos$value[index_val] + channel_stacks$orders[i]
}

#Populate the conversion node values
display_node_labels = node_labels
for(i in 1:length(label_length)){
  for(j in 1:label_length[i]){
    display_node_labels[j+ifelse(i==1,0,sum(label_length[1:(i-1)]))] = paste0(i,":",node_labels[j+ifelse(i==1,0,sum(label_length[1:(i-1)]))])
  }
}
display_node_labels = c(display_node_labels, "Conversion")

```

# Building the Plot

And finally, I can generate the actual Sankey diagram:

```{r}
#Generate Sankey diagram
p <- plot_ly(
    type = "sankey",
    orientation = "v",

    node = list(
      label = display_node_labels,
      #color = node_colors,
      pad = 10,
      thickness = 30,
      line = list(
        color = "white",
        width = 0
      )
    ),
  
    link = list(
      source = combos$source-1, # convert to zero index
      target = combos$target-1, # convert to zero index
      value = combos$value, #size of connection
      color = "rgba(0, 0, 0, 0.2)"
    )
  ) %>% 
  layout(
    title = "Conversion Flow Diagram",
    font = list(
    size = 10
    )
  )
p
```




















