---
title: "Load Clusters Into Workspace"
author: "Trevor Paulsen"
output: html_notebook
---

# Convert Data Feed Visitor IDs to Hex
The first step is translating the visitor IDs into a format that customer attributes will understand.

```{r}
library(Rmpfr)
library(RCurl)

example_vis_ids = as.data.frame(
  c(
    "3012860238288533203_4611687811039376400", 
    "3012860276943234879_6917530783211035360", 
    "3012860334925186994_6917530727911358233"
    )
  )
names(example_vis_ids) = "visids"

# Convert the integer visitor IDs to hex visitor IDs
convert_to_hex = function(visitor_id){
  visid_hi = toupper(as.character(formatMpfr(mpfr(strsplit(visitor_id,"_")[[1]][1], base=10), base=16, decimal.mark="")))
  visid_lo = toupper(as.character(formatMpfr(mpfr(strsplit(visitor_id,"_")[[1]][2], base=10), base=16, decimal.mark="")))
  visitor_id_hex = paste0(visid_hi,"-",visid_lo)
  return(visitor_id_hex)
}

hex_visitor_ids = example_vis_ids %>%
  mutate(
    hex_visids = convert_to_hex(as.character(visids))
  )

hex_visitor_ids
```


# Save off file for Customer Attribute Import
Next, we'll save a file that can be imported to the Customer Attributes feature.
```{r}
# Note that this technique only works on the cookie based ID, not the "stitched ID"

# Combine clusters with hex values we created
cluster_mapping = data.frame(hex_visitor_ids, export_table)
cluster_mapping = data.frame(visitor_ids_converted_example$`Hex ID`, export_table$cluster)
names(cluster_mapping) = c("aaid", "cluster")

# Write the lookup to a file
write.table(cluster_mapping, file="cluster_mapping.csv", sep=",", row.names = FALSE, quot=FALSE)

# Prebaked Example
cluster_mapping
```

# Fire off "setCustomerID" a bunch of times
Now we'll actually load the data into Analytics by spoofing a setCustomerID API call.
I've heard there's a way to do this via batch ingest, but it's not documented...

What's cool about this is you don't even have to have MCID implemented for this to work!

```{r}
# setCustomerID a zillion times (try doing this in parrellel)

org_id = "<YOUR IMS ORG>@AdobeOrg"
for (i in 1:3){
  url = paste0("https://dpm.demdex.net/id?",
               "d_visid_ver=1.8.0&",
               "d_fieldgroup=AAM&",
               "d_rtbd=json&",
               "d_ver=2&",
               "d_orgid=", org_id, "&",
               "d_nsid=0&",
               "d_blob=NRX38WO0n5BH8Th-nqAG_A&",
               "d_cid_ic=AVID%01", hex_visitor_ids$hex_visids[i], "&",
               "d_cid_ic=aaid%01", hex_visitor_ids$hex_visids[i], "&",
               "ts=1491014875614")

  response = getURL(url = url, ssl.verifyhost = FALSE, ssl.verifypeer = FALSE, followlocation = TRUE)
  response
}
```



